### 2020-05-07 paper/midway.tex
Midway board paper is underway!

\documentclass[12pt, twocolumn]{article}
\usepackage{cite}

\author{Sammy Furr}
\title{Teaching Debugging Collaboratively---Midway Report}
\date{\today}

\begin{document}

\begin{titlepage}
\maketitle
\end{titlepage}

\section{Introduction}

Debugging is invaluable in writing and understanding code, yet it is
rarely formally taught.  We typically teach students programming
structures, concepts, and languages, but we leave them to learn the
tools they use to code by themselves.  This approach often works
well---the programmer's choice of editor is \textit{very} personal,
students figure out how to configure an individualized workflow.
Perhaps because debuggers are tools, they often get lumped into the
``teach yourself'' category.  Unlike editors or reference guides
however, effectively using a debugger requires a set of high-level,
platform agnostic, teachable skills.  Teaching these skills is
effective, and translates into better, faster, debugging and
programming.\cite{10.1145/3286960.3286970}\cite{10.1145/3361721.3361724}\par

Teaching debugging is particuarly important in

\section{The Value of Teaching Debugging}

\section{Debugger: Low Level}

\section{Debugger: High Level}

\section{Next Steps}

\pagebreak
\bibliographystyle{acm}
\bibliography{sprojbib}{}
\end{document}